{\bf \huge Wikiplia}:\\ The Free Programming Language\\ That Anyone Can Edit

### 2007-03-20 trunk/wikiplia/paper/wikiplia.tex
interface ready for paper

% TODO: license
\newcommand\currentrevision{465}

@@

The revision control system stores XESP documents and their history.
This is mostly a straightforward implementation of the imperative {\tt
cvs} judgments given in Section~\ref{sec:calculus}. Unlike typical
revision control systems, we need to support very large numbers of
revisions with small edits (since every change is saved), so the
implementation is engineered to make storing small revisions very
cheap. Particularly, revisions are aggressively compressed by only
storing the newest revision directly, and then a series of difference
``plans'' that describe how to get the next older revision from the
current one. We compute optimal plans using an efficient minimal edit
distance\wcite{Levenshtein distance} calculation at the
token\wcite{Token (parser)} level. As of revision~\currentrevision,
the database is only 236 kilobytes.\z\wcite{Kilobyte}

@@

Though Wikipedia in its current form is a usable general-purpose
programming language, work remains to be done for it to reach its full
potential. For one, it needs a vibrant community of contentious and
hubristic editors hiding behind anonymous IP addresses boldly
asserting half-baked syntactic extensions or enforcing superficial
style preferences, gritting their teeth while typing and clicking
white knuckled in a kind of road rage\wcite{List of rages} created by
the dehumanizing semantic markup by which they are forced to
communicate.

We also seek to improve the languages. The language {\tt w} needs many
more features to speed development: the parenthesis-based XESP syntax
should eventually be replaced by a pleasant concrete syntax, if we can
get around to it before too much code is written in XESP.

@@

All popular modern languages are defined via a definitional
interpreter\wcite{JavaScript}$^,$\wcite{Objective
Caml}$^,$\wcite{Perl} with accompanying O'Reilly ``animal''
book.\z\wcite{O'Reilly Media} The work on Wikiplia is unrelated: We
have no animal mascot\wcite{ORLY owl} and the languages are described
by a tower of source-to-source translations on top of a universally
parseable semantic document in XML\wcite{Category:ML programming
language family} form.

The author doesn't think much of musicals,\z\wcite{Musical theatre} to
be perfectly honest, so those are basically a no-go. He also feels
that the metric system\wcite{Metric system} and time zones\wcite{Time
zone} are pretty overrated. Ketchup on eggs is gross, but not quite as
gross as foie gras\wcite{Foie gras}, which more or less has the word
``gross'' in its name so duh.\z\wcite{Duh}

@@

We have described Wikiplia, the free programming language that anyone
can edit. Unlike other programming languages, it supports a variety of
freedoms (...) and is explicitly scalable to new freedoms. ...

>>>>

\newcommand\currentrevision{532}
\newcommand\currentsizekb{250}
\newcommand\wikipliaurl{http://wikiplia.spacebar.org/}

@@

The revision control system stores XESP documents and their history.
This is mostly a straightforward implementation of the imperative {\tt
cvs} judgments given in Section~\ref{sec:calculus}. Unlike typical
revision control systems, we need to support very large
numbers\footnote{The implementation also supports revision numbers of
arbitrary size.} of revisions with small edits (since every change is
saved), so the implementation is engineered to make storing small
revisions very cheap. Particularly, revisions are aggressively
compressed by only storing the newest revision directly, and then a
series of difference ``plans'' that describe how to get the next older
revision from the current one. We compute optimal plans using an
efficient minimal edit distance\wcite{Levenshtein distance}
calculation at the token\wcite{Token (parser)} level. As of
revision~\currentrevision, the database is only \currentsizekb\ 
kilobytes.\z\wcite{Kilobyte}

@@

Though Wikiplia in its current form is a usable general-purpose
programming language, work remains to be done for it to reach its full
potential. For one, it needs a vibrant community of contentious and
hubristic editors hiding behind anonymous IP addresses boldly
asserting half-baked syntactic extensions or enforcing superficial
style preferences, gritting their teeth while typing and clicking
white knuckled in a kind of road rage\wcite{List of rages} created by
the dehumanizing semantic markup by which they are forced to
communicate.

% XXX: supporting users with revision #s?

We also seek to improve the languages. The language {\tt w} needs many
more features to speed development: the parenthesis-based XESP syntax
should eventually be replaced by a pleasant concrete syntax, if we can
get around to it before too much code is written in XESP. A type
system\wcite{Type system} is not planned, because type systems
restrict Freedom~\fexpression, the freedom of expression. However, we
should seek to make Wikiplia as multi-paradigm as possible (again,
freedom from discrimination on the basis of paradigm\wcite{paradigm}
orientation), supporting OOPs-oriented
programming,\z\wcite{Object-oriented programming} aspect-oriented
programming,\z\wcite{Aspect-oriented programming} duck-oriented
typing,\z\wcite{Duck typing} orientation-oriented
orienteering,\z\wcite{Orienteering} Orient-oriented
programming,\z\wcite{The Orient} etc.

The {\tt page} language needs extensions for developing human-readable
web pages, mostly for the purpose of creating jazzy graphics and boxes
that distract from or directly call attention to obvious problems with
the pages without actually addressing those problems.

@@

All popular modern languages are defined via a definitional
interpreter\wcite{JavaScript}$^,$\wcite{Objective
Caml}$^,$\wcite{Perl} with accompanying O'Reilly ``animal''
book.\z\wcite{O'Reilly Media} The work on Wikiplia is unrelated: We
have no animal mascot\wcite{ORLY owl} and the languages are described
by a tower of source-to-source translations on top of a universally
parseable semantic document in XML\wciteshort{Category:ML programming
language family} form.

The author doesn't think\wcite{Thought} much of
musicals,\z\wcite{Musical theatre} to be perfectly\wcite{Thomas
Aquinas} honest,\z\wcite{Honesty} so those are basically\wcite{BASIC}
a no-go. He also feels that the metric system\wcite{Metric system} but
paradoxically\wcite{Pardadox} also time zones\wcite{Time zone} are
pretty overrated. Ketchup\wcite{Ketchup} on eggs\wcite{Egg (food)} is
gross,\z\wcite{Gross} but not quite as gross as foie gras\wcite{Foie
gras}, which more or less has the word\wcite{Word} ``gross'' in its
name\wcite{Name} so duh.\z\wcite{Duh}

@@

We have described Wikiplia, the free programming language that anyone
can edit. Unlike other programming languages, it is designed to
support a variety of freedoms (\ftinker, \fexpression, \fbeer, \fusa,
\ffree, \ftobe) and is explicitly scalable to new freedoms. The system
is implemented in a minimal bootstrapping core, and then built up to
a featured system using its own faculties for extension. XXX more...

\bigskip
$${\tt \wikipliaurl}$$
